Power and Pitfalls of Generic Smart Contracts

### 2021-07-29 001_abstract.tex
Added omments on our solution

Generics are a powerful feature of programming languages that allows one
to write highly reusable code.
%
More specifically, they are based on the use of type placeholders in order
to produce parametrized code, that can be instantiated for each
concrete type provided for them. 
%
In many programming languages, such as Java, they are implemented by
\emph{erasure}\ie replaced by their upper bound type during bytecode compilation.
%
This paper investigates the use of generics in the
context of smart contracts for blockchain, in order to implement
a contract for \emph{shared entities} (such as a company shared
by its shareholders),
by using the Hotmoka blockchain whose contracts are written in Java.
%
The considered case study is particular important since
a special case of shared entities could be the validators' set of
the blockchain itself. The analysis shows that, the power of generics comes at the risk of a too permissive
typing of the compiled code, due to the erasure mechanism, with a consequent possible attack
to the validators' set. This paper proposes a solution
that forces the compiler to generate more precise type information than
those arising by erasure.

>>>>

Generics are a powerful feature of programming languages that allows one
to write highly reusable code.
%
More specifically, they are based on the use of type placeholders in order
to produce parametrized code, that can be instantiated for each
concrete type provided for them. 
%
In many programming languages, such as Java, they are implemented by
\emph{erasure}\ie replaced by their upper bound type during compilation into bytecode.
%
This paper investigates the use of generics in the
context of smart contracts for blockchain, in order to implement
a contract for \emph{shared entities} (such as a company shared
by its shareholders),
by using the Hotmoka blockchain whose contracts are written in Java.
%
The considered case study is particularly important since
a special case of shared entities could be the validators' set of
the blockchain itself. The analysis shows that the power of generics comes at the risk of a too permissive
typing of the compiled code, due to the erasure mechanism, with a consequent possible attack
to the validators' set. This paper proposes a solution
that forces the compiler to generate more precise type information than
those arising by erasure.

### 2021-07-29 040_solution.tex
Added omments on our solution

\todo{Si potrebbe dire che la soluzione e' simile ad una gestione dei generics eterogenea? che pero' deve essere fatta a mano ridefinendo i metodi, invece che essere generata in automatico dal compilatore? }

>>>>

The solution of redefining method \<accept> can be seen as a limited form of heterogeneous
compilation of generics, restricted to a specific method and forced manually, rather than
automatically performed by the compiler. It is interesting to consider which methods
would need that redefinition, in general. They are those that have a parameter of generic type and
that generic type is restricted in the subclass. For instance, method \<accept> in
Fig.~\ref{fig:shared_entity} has parameters \<buyer> and \<offer> of generic type
\<S> and \<O>, respectively. The subclass in Fig.~\ref{fig:abstract_validators} restricts
\<S> to be a \<TendermintED25519Validator> and \<O> to be
an {\codesize\texttt{Offer<TendermintED25519Validator>}}. Hence we must redefine
\<accept> in the subclass with the more specific types for the \<buyer> and \<offer>
parameters. In general, a compiler might perform this automatically or a static analysis
tool might issue a warning when such redefinition is needed. Currently, however, this is
left to the programmer, which might be the origin of security issues, as shown
in Sec.~\ref{sec:attack}.
